---
title: "Untitled"
author: "Katie Lotterhos"
date: "6/27/2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("0-Novelty_Oceans_Functions.R")
```


setwd("~/Desktop/Repos/OceanClimateNovelty/src")

```{r load data}
dat_Nov <- fread("../results/SigmaD.csv")
dat4.5 <- fread("../data/large_files/T_Ar_Ca_pH_RCP45.txt", sep = ",")
dat8.5 <- fread("../data/large_files/T_Ar_Ca_pH_RCP85.txt", sep = ",")
```

In the column names, the first one is the baseline (A) and the second one corresponds
to the station (B). 

When the future (B) is queried against the past baseline (A), it represents novel climates in the future. 

When the past (B) is queried against the future baseline (A), it represents disappearing climates in the past.

NN.sigma represents Mahoney's $\sigma_D$

However is some cases our Mahalanobis distances were so extreme they gave the 
maximum possible value of $\sigma_D$. So I will compare `Mdist` to see if they
represent the data better.

* Upper left: Disappearing climates from 1800 in today's ocean (NN.sigma_today_1800)
* Upper right: Novel climates today from 1800 (NN.sigma_1800_today)
* Middle left: Disappearing climates from today under RCP 4.5 at 2100 (NN.sigma_2100_4.5_today)
* Middle right: Novel climates for RCP 4.5 from today (NN.sigma_today_2100_4.5)
* Lower left: Disappearing climates from today under RCP 8.5 at 2100 (NN.sigma_2100_8.5_today)
* Lower right: Novel climates for RCP 8.5 from today (NN.sigma_today_2100_8.5)

```{r}
B <- seq(0,8, by=0.01)



pA <- Plot_interp(dat_Nov, dat_Nov$NN.sigma_today_1800, B , "A) 1800 climates disappearing in 2000") #Disappearing from 1800
pB <- Plot_interp(dat_Nov, dat_Nov$NN.sigma_1800_today, B, "B) Novel climates in 2000 since 1800") #Novel today from 1800

pC <- Plot_interp(dat_Nov, dat_Nov$NN.sigma_2100_4.5_today, B, "C) 2000 climates disappearing in 2100 RCP 4.5") #Disappearing from today
pD <- Plot_interp(dat_Nov, dat_Nov$NN.sigma_today_2100_4.5, B, "D) Novel climates in 2100 RCP 4.5 since 2000") #Novel 2100 4.5

pE <- Plot_interp(dat_Nov, dat_Nov$NN.sigma_2100_8.5_today,B,  "E) 2000 climates disappearing in 2100 RCP 8.5") #Disappearing from today
pF <- Plot_interp(dat_Nov, dat_Nov$NN.sigma_today_2100_8.5, B, "F) Novel climates in 2100 RCP 8.5 since 2000") #Novel 2100 8.5

p_legend <- Plot_interp_legend(dat_Nov, dat_Nov$NN.sigma_today_2100_8.5, 0:8, "Sigma\ndissimilarity")

pdf("../results/SigmaDmaps.pdf", width=11, height=7)
  grid.newpage()
  page.layout <- grid.layout(nrow=3, ncol= 3, widths = c(0.43,0.43, 0.14), heights=c(1/3, 1/3, 1/3))
  pushViewport(viewport(layout=page.layout))
  
  print(pA, vp=viewport(layout.pos.row=1, layout.pos.col = 1))
  print(pB, vp=viewport(layout.pos.row=1, layout.pos.col = 2))
  print(pC, vp=viewport(layout.pos.row=2, layout.pos.col = 1))
  print(pD, vp=viewport(layout.pos.row=2, layout.pos.col = 2))
  print(pE, vp=viewport(layout.pos.row=3, layout.pos.col = 1))
  print(pF, vp=viewport(layout.pos.row=3, layout.pos.col = 2))
  
  print(p_legend, vp=viewport(layout.pos.row=2, layout.pos.col = 3))
dev.off()



hist(dat_Nov$NN.sigma_today_2100_8.5)
hist(dat_Nov$NN.sigma_2100_8.5_today)

image(dat_Nov$long, dat_Nov$lat, dat_Nov$)

```