---
title: "Untitled"
author: "Katie Lotterhos"
date: "6/27/2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("0-Novelty_Oceans_Functions.R")
```


setwd("~/Desktop/Repos/OceanClimateNovelty/src")

```{r load data}
dat_Nov <- fread("../results/SigmaD.csv")
dat4.5 <- fread("../data/large_files/T_Ar_Ca_pH_RCP45.txt", sep = ",")
dat8.5 <- fread("../data/large_files/T_Ar_Ca_pH_RCP85.txt", sep = ",")
```

In the column names, the first one is the baseline (A) and the second one corresponds
to the station (B). 

When the future (B) is queried against the past baseline (A), it represents novel climates in the future. 

When the past (B) is queried against the future baseline (A), it represents disappearing climates in the past.

NN.sigma represents Mahoney's $\sigma_D$

However is some cases our Mahalanobis distances were so extreme they gave the 
maximum possible value of $\sigma_D$. So I will compare `Mdist` to see if they
represent the data better.

* Upper left: Disappearing climates from 1800 in today's ocean (NN.sigma_today_1800)
* Upper right: Novel climates today from 1800 (NN.sigma_1800_today)
* Middle left: Disappearing climates from today under RCP 4.5 at 2100 (NN.sigma_2100_4.5_today)
* Middle right: Novel climates for RCP 4.5 from today (NN.sigma_today_2100_4.5)
* Lower left: Disappearing climates from today under RCP 8.5 at 2100 (NN.sigma_2100_8.5_today)
* Lower right: Novel climates for RCP 8.5 from today (NN.sigma_today_2100_8.5)

```{r}
Plot_interp(dat_Nov, dat_Nov$NN.sigma_today_1800, "Yo") #Disappearing from 1800
Plot_interp(dat_Nov, dat_Nov$NN.sigma_1800_today, "Yo") #Novel today from 1800

Plot_interp(dat_Nov, dat_Nov$NN.sigma_2100_4.5_today, "Yo") #Disappearing from today
Plot_interp(dat_Nov, dat_Nov$NN.sigma_today_2100_4.5, "Yo") #Novel 2100 4.5

Plot_interp(dat_Nov, dat_Nov$NN.sigma_2100_8.5_today, "Yo") #Disappearing from today
Plot_interp(dat_Nov, dat_Nov$NN.sigma_today_2100_8.5, "Yo") #Novel 2100 8.5

hist(dat_Nov$NN.sigma_today_2100_8.5)
hist(dat_Nov$NN.sigma_2100_8.5_today)

image(dat_Nov$long, dat_Nov$lat, dat_Nov$)

```